---
title: UNIT08B：多元尺度分析(MDS)與品牌知覺圖
author: 中山大學管理學院 卓雍然
date: "`r Sys.time()`"
output: 
  html_document:
    highlight: pygments
    theme: flatly
    css: ../style.css
---

```{r results='hide', message=FALSE, warning=FALSE, echo=F}
# This Chunk is for format setting, Do NOT Change the codes!
rm(list=ls(all=T))
knitr::opts_chunk$set(comment = NA)
knitr::opts_knit$set(global.par = TRUE)
par(cex=0.8)
options(scipen=20, digits=5, width=80)
if(!require(pacman)) install.packages("pacman")
```
<hr>

<p class="wwl">
<span style="font-size:24px">`r "\U1F4A1"` 學習重點：</span><br>
&emsp; ■ 多元尺度分析(MDS, Multi-Dimension Scaling)<br>
&emsp; &emsp; § 縮減尺度工具<br>
&emsp; &emsp; § 盡量維持持高、低尺度之中各資料點間的距離<br>
&emsp; ■ 品牌知覺圖(Perception Map)：呈現在消費者心中<br>
&emsp; &emsp; § 品牌與品牌間的關係<br>
&emsp; &emsp; § 屬性與屬性間的關係<br>
&emsp; &emsp; § 品牌與屬性之間的關係<br>
</p class="wwl">

<br>
```{r results='hide', message=FALSE, warning=FALSE}
pacman::p_load(dplyr, FactoMineR, factoextra)
```

##### § Reading and Preapring Data
```{r}
cafe = read.csv("data/cafe2.csv")
row.names(cafe) = cafe[,1]
cafe = cafe[,-1]
cafe
```
這裡我們有5個品牌(資料點)、8種屬性(尺度)，儘管資料點數不多，我們還是很難同時想像八個尺度
<br>

### 1. Multi-Dimensional Scaling (多元尺度分析)

+ Calculate the distance matrix by `dist()`
+ Call `cmdscale()` for MDS analysis.

```{r}
d = dist(cafe, method = "euclidean")
fit <- cmdscale(d, eig = TRUE, k = 2)
```

+ Then plot the brands on the space of reduced dimensions. 

```{r fig.height=6, fig.width=5.5}
brands = rownames(cafe); percepts=colnames(cafe)
cafe$x = fit$points[, 1]
cafe$y = fit$points[, 2]
plot(cafe$x, cafe$y, pch = 19, xlim=c(-1.5,1.5), ylim=c(-1,1.5))
bpos = c(3, 3, 4, 4, 3)
text(cafe$x, cafe$y, pos=bpos, offset=0.5, labels = brands)
```

Perceived differennce among the brands are represented as the corresponding distances on the map.
<br>

### 2. Percpetual Map (知覺圖)
Furthermore, we can also plot the unit vectors for the original dimension on the map.
```{r fig.height=6, fig.width=5.5}
plot(cafe$x, cafe$y, pch = 19, xlim=c(-3.5,2), ylim=c(-2.2,2.5))
for(i in 1:length(percepts)) {
  m = lm(sprintf("%s ~ x + y",percepts[i]), data=cafe)
  c = 1/m$coef[2]; d = 1/m$coef[3]
  cd2 = c*c + d*d
  px = (c * d * d)/cd2
  py = (c * c * d)/cd2
  arrows(0,0,px,py,col="pink",length=0)
  text(px, py, col="red", cex=0.75, labels = percepts[i])
}
text(cafe$x, cafe$y, pos=bpos, offset=0.5, cex=0.75, labels = brands)
```

### 3. 主成分分析 PCA
```{r fig.width=6, fig.height=5}
PCA(cafe[1:8], graph=F) %>% fviz_pca_biplot(
  repel=T, col.var="red", col.ind="blue", 
  labelsize=3, pointshape=16, pointsize=3)
```
<br><hr>

<p class="qiz">
<span style="font-size:24px">`r "\U1F5FF"` 問題討論：</span><br>
&emsp; ■ 請你比較「主成分分析(PCA)」和「多元尺度分析(MDS)」的異同？<br>
&emsp; ■ 請你從品牌知覺圖之中觀察到哪一些現象？<br>
</p class="qiz"><br>

